Carbonate chemistry at Ny-Ålesund

### 2022-08-16 awipev-CO2.Rmd
In stratification: fCO2 replaced by pCO2

# we look at pCO2  at all depths
# calculate fCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_sf = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    fco2_sf = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_sf,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$fCO2
  ) %>% 
  dplyr::select(datetime, fco2_sf)
d_all <- left_join(d_all, tmp, by = "datetime")
tmp <- d_all %>% 
  dplyr::select(montht, fco2_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = fco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "fco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$fco2_sf ~ tmp$depth_group * tmp$montht))

# we look at fCO2 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, fco2_sf, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$fco2_sf ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(fco2_sf_med = median(fco2_sf, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = fco2_sf_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   dplyr::mutate(fco2_sf_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet fCO2 with (median0_4m - median8_12m)")

p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group2), y = fco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pH at 4°C vs depth", x = "Depth (m)", y= "pHint_tot_sf") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

fco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = fco2_sf_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(substitute("fCO"["2"]), limits = c(-45, 10)) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 2.5,
    y = 5,
    parse = TRUE,
    label = "fCO['2 surface'] > fCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 2.5,
    y = -40,
    parse = TRUE,
    label = "fCO['2 surface'] < fCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
fco2_delta_plt
# all deltas
deltas_plt <- sal_delta_plt / temp_delta_plt / den_delta_plt / fco2_delta_plt
deltas_plt <- deltas_plt + plot_annotation(tag_levels = 'a')
print(deltas_plt)
ggsave(filename = "figures/deltas.png", plot = deltas_plt, width = 15, height = 20, units = "cm")
```

>>>>

# we look at pCO2  at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_sf = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_sf = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_sf,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2
  ) %>% 
  dplyr::select(datetime, pco2_sf)
d_all <- left_join(d_all, tmp, by = "datetime")
tmp <- d_all %>% 
  dplyr::select(montht, pco2_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_sf ~ tmp$depth_group * tmp$montht))

# we look at pCO2 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, pco2_sf, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pco2_sf ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(pco2_sf_med = median(pco2_sf, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pco2_sf_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   dplyr::mutate(pco2_sf_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet fCO2 with (median0_4m - median8_12m)")

p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group2), y = pco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pCO2", x = "Depth (m)", y= "pHint_tot_sf") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

pco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_sf_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(substitute("pCO"["2"]), limits = c(-45, 10)) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 2.5,
    y = 5,
    parse = TRUE,
    label = "pCO['2 surface'] > pCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 2.5,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface'] < pCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
pco2_delta_plt
# all deltas
deltas_plt <- sal_delta_plt / temp_delta_plt / den_delta_plt / pco2_delta_plt
deltas_plt <- deltas_plt + plot_annotation(tag_levels = 'a')
print(deltas_plt)
ggsave(filename = "figures/deltas.png", plot = deltas_plt, width = 15, height = 20, units = "cm")
```
